Machine Learning Models for the Segmentation of Clinical Exchange

### 2018-03-08 paper-amia-2018/manuscript.tex
update results section

 NB & 0.594 & 0.662 & 0.626 & 0.590 & 0.666 & 0.626 \\ \hline
 SVM & 0.742 & \textbf{0.679} & 0.709 & 0.774 & 0.696 & 0.733\\ \hline
 KNN & \textbf{0.808} & 0.663 & \textbf{0.728} & \textbf{0.820} & \textbf{0.742} & \textbf{0.779}\\ \hline
 LSTM & -- & -- & -- & 0.800 & 0.646 & 0.714  \\ \hline
 GRU & -- & -- & -- & 0.769 & 0.715 & 0.741 \\ \hline 
  \end{tabular}
\end{table}

As follows from Table~\ref{tab:result_boundary}, KNN performs best among all machine learning models in terms of precision and F1-measure, achieved 0.808 precision with 0.728 F1-measure when contextual features are used, and 0.820 precision with 0.779 F1-measure when a combination of contextual, topic, and punctuation mark features are used. However, RNN demonstrates the lowest performance among all models in terms of recall and F1-measure while GRU shows 3.72\%, 17.79\% and 11.47\% higher precision, recall, and F1-measure than LSTM. In this study, SVM appears as the second highest model in terms of precision and F1-measure, obtains highest 0.679 recall when only textual features are used. On the other hand, NB exhibits lowest precision value 0.594, provides better recall and F1-measure than RNN model. When textual features are used in combination with topic and punctuation mark features, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively. Nevertheless, precision increases by 4.31\% and 1.49\% for SVM and KNN methods while decreases by 0.7\% in NB. \\

@@

Table~\ref{tab:result_not_boundary} summarizes the performance of NB, SVM, KNN, and RNN models for detecting ``not boundary'' class in e-coaching text. We observed that performance is remarkably high in ``not boundary'' class compared to boundary detection which is expected because 96.71\% instances are from ``not boundary'' class. Similar to boundary detection, KNN consistently outperforms over all other methods but obtains 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure for contextual features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure for combined features compared to boundary class. However, RNN demonstrates the lowest performance and exhibits 0.981 and 0.983 precision with 0.986 and 0.987 F1-measure for LSTM and GRU, respectively. Impact of additional features is also consistent with ``not boundary'' classification. Results show that F1-measure increases by 0\%, and 0.1\% for SVM and KNN models although decreases by 0.1\% for NB. \\

@@

Table~\ref{tab:result_weighted_avg} outlines the weighted average results of the experiment on the models for the segmentation of e-coaching text by classifying them into ``boundary'' and ``not boundary'' classes. Overall, KNN obtains the best performance with all metrics and RNN denotes the lowest performance among all methods. NB and SVM demonstrate moderate performance, obtain precision 0.975 and 0.981; recall 0.974 and 0.982; and F1-measure 0.975 and 0.983 when textual features are used. Influence of the additional features is also consistent as above, precision increases by 0.1\%, 0.2\%, and 0.3\%; recall increases by 0\%, 0.1\%, and 0.2\%; and F1-measure increases by 0\%, 0.2\%, and 0.3\% for NB, SVM, and KNN methods, respectively, when combined features are used.\\

@@

We observed second highest performance of RNN, in particular, LSTM and GRU for the text segmentation. We believe that RNN will perform better if a large set of data is utilized. In this study, we employed 3,138 examples of boundary case which failed to achieve best results. We also observed that GRU performs better than LSTM which was observed in other previous study\cite{chung2014empirical}.

>>>>

 NB & 0.594 & 0.662 & 0.626 & 0.590 & 0.666 & 0.626 \\ \hline
 SVM & 0.742 & \textbf{0.679} & 0.709 & 0.774 & 0.696 & 0.733\\ \hline
 KNN & \textbf{0.808} & 0.663 & \textbf{0.728} & \textbf{0.820} & \textbf{0.742} & \textbf{0.779}\\ \hline
 LSTM & -- & -- & -- & 0.800 & 0.646 & 0.714  \\ \hline
 GRU & -- & -- & -- & 0.800 & 0.715 & 0.741 \\ \hline 
  \end{tabular}
\end{table}

As follows from Table~\ref{tab:result_boundary}, KNN performs best among all machine learning models in terms of precision and F1-measure, achieved 0.808 precision with 0.728 F1-measure when contextual features are used, and 0.820 precision with 0.779 F1-measure when a combination of contextual, topic, and punctuation mark features are used. However, NB demonstrates the lowest performance among all models in terms of all performance metrics. In this study, GRU appears as the second highest model, obtains 10.68\% higher recall with 3.78\%  higher F1-measure than LSTM. On the other hand, SVM exhibits highest recall 0.679 when only textual features are used. When textual features are used in combination with topic and punctuation mark features, recall increases by 0.6\%, 2.5\%, and 11.92\%; and F1-measure increases by 0\%, 3.39\%, and 7\% for NB, SVM, and KNN models, respectively. Nevertheless, precision increases by 4.31\% and 1.49\% for SVM and KNN methods while decreases by 0.7\% in NB. \\

@@

Table~\ref{tab:result_not_boundary} summarizes the performance of NB, SVM, KNN, and RNN models for detecting ``not boundary'' class in e-coaching text. We observed that performance is remarkably high in ``not boundary'' class compared to boundary detection which is expected because 96.71\% instances are from ``not boundary'' class. KNN achieves 22.40\%, 50.07\%, and 36.26\% higher precision, recall, and F1-measure for contextual features; and 20.85\%, 33.96\%, and 27.47\% higher precision, recall, and F1-measure for combined features compared to boundary class. In contrast to boundary detection, RNN demonstrates the highest performance among all models. LSTM obtains 0.993 precision with 0.994 recall and F1-measure while LSTM exhibits 0.994 for all performance metrics. Impact of additional features is also consistent with ``not boundary'' classification. Results show that F1-measure increases by 0\%, and 0.1\% for SVM and KNN models although decreases by 0.1\% for NB. \\

@@

Table~\ref{tab:result_weighted_avg} outlines the weighted average results of the experiment on the models for the segmentation of e-coaching text by classifying them into ``boundary'' and ``not boundary'' classes. Overall, KNN obtains the best performance with all metrics and NB denotes the lowest performance among all methods. GRU demonstrates the same result as KNN for precision and F1-measure except for recall. SVM shows moderate performance, obtains 0.981, 0.982, and 0.983 for precision, recall, and F1-measure, respectively when textual features are used. Influence of the additional features is also consistent as above, precision increases by 0.1\%, 0.2\%, and 0.3\%; recall increases by 0\%, 0.1\%, and 0.2\%; and F1-measure increases by 0\%, 0.2\%, and 0.3\% for NB, SVM, and KNN methods, respectively, when combined features are used.\\

@@

We observed the second highest performance of RNN, in particular, LSTM and GRU for the text segmentation. We believe that RNN will perform better if a large set of data is utilized. In this study, we employed 3,138 examples of boundary case which failed to achieve best results. We also observed that GRU performs better than LSTM which was observed in other previous study\cite{chung2014empirical}.
